Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-02 parallel-tractability-J-v9/evaluation.tex
worked on experimental section, fixed table layout and captions

\subsection{Practical Usability of the Theoretical Results}

In this part, we analyze different kinds of ontologies and datasets
including benchmarks, real-world ontologies and
datasets that can be expressed in ontology languages.
Based on the analysis of these datasets,
we find that, ignoring imports, many of them
belong to $\mathcal{D}_{\textit{\text{dhl}}}$ or
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.

\textbf{Benchmarks}. In the
Semantic Web community, many benchmarks are proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way. We investigate several
popular benchmarks using our results. For some benchmarks, the ontologies have TBoxes with a simple
structure. These ontologies can be expressed in RDFS
and belong to $\mathcal{D}_{\textit{\text{dhl}}}$.
These benchmarks include SIB\footnote{https://www.w3.org/wiki/Social$\underline{~}$Network$\underline{~}$Intelligence$\underline{~}$BenchMark}
(\emph{Social Network Intelligence BenchMark}),
BSBM\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/berlinsparqlbenchmark/}
(\emph{Berlin SPARQL Benchmark}) and
LODIB\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/lodib/}
(\emph{Linked Open Data Integration Benchmark}).
The ontology used in IIMB\footnote{http://islab.di.unimi.it/iimb/}
(\emph{The ISLab Instance Matching Benchmark}) satisfies the simple-concept restriction.

In the latest version of LUBM\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
(The Lehigh University Benchmark),
there are 48 classes and 32 properties.
Statements about properties, such as inverse property statements,
can be rewritten into datalog rules allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Most of the statements about classes can be rewritten into datalog rules that
are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$,
where a new anonymous constant $y$ is introduced.
This kind of rule is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
On the other hand, in some cases, this kind of rule can also be
eliminated when taking a rewriting approach \cite{GrauHKKMMW13}.
In summary, if the above kind of rule is not considered,
the materialization of a LUBM dataset can be handled by Algorithm~$\mathsf{A}_{\text{opt}}^{\psi}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet and the latest version
YAGO3 \cite{MahdisoltaniBS15} has more than 10 million entities
(e.g., persons, organizations, cities, etc.)
and contains more than 120 million facts about these entities.
In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called YAGO \emph{model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}.
In addition to the expressiveness of RDFS,
YAGO \emph{model} also allows for stating the \emph{transitivity} and \emph{acyclicity} of a property.
Making full use of RDFS features cannot lead to parallel tractability \cite{Horst05}.
However, \citet{SuchanekKW08} specify a more efficient set of materialization rules.
All of these rules are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Thus, we have that a well-constructed YAGO dataset belongs to $\mathcal{D}_{\textit{\text{dhl}}}$.

\textbf{Real Ontologies}. We investigated
$151$ ontologies that cover many domains
like biomedicine, geography, etc.
These ontologies are collected
from the Protege ontology library,\footnote{http://protegewiki.stanford.edu/wiki/Protege$\underline{~}$Ontology$\underline{~}$Library}
Swoogle\footnote{http://swoogle.umbc.edu/}
and the Oxford ontology library.\footnote{http://www.cs.ox.ac.uk/isg/ontologies/lib/}
All ontologies are available online.\footnote{https://github.com/quanzz/PT}
Among these ontologies, $111$ belong to $\mathcal{D}_{\textit{\text{dhl}}}$
or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$; $21$ of the DHL ontologies contain
conjunctions and satisfy the simple-concept restriction.
The remaining ontologies have simple TBoxes, i.e., no conjunction ($A_1\sqcap A_2$)
appears in these ontologies.
We also find two DHL($\circ$) ontologies and both satisfy the simple-role restriction.

The above investigation indicates that the simple-concept and the simple-role restrictions
is naturally satisfied in many real applications of different fields.
From the perspective of developers who work on building their own ontologies,
they can also refer to $\mathcal{D}_{\textit{\text{dhl}}}$ and $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
to achieve the guarantee of the parallel tractability.

@@

\begin{figure}[htbp]
  \centering
  \begin{minipage}{.24\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/1-Finance-simple}
    \subcaption{lg1\label{fig:financepdhl}} 
  \end{minipage}
  \begin{minipage}{.24\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/2-molecule-simple}
    \subcaption{lg3\label{fig:moleculepdhl}}
  \end{minipage}
  \begin{minipage}{.24\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/3-NIF_GrossAnatomy-simple}
    \subcaption{lg5\label{fig:grossanatomypdhl}}
  \end{minipage}
  \begin{minipage}{.24\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/4-skeleton-simple}
    \subcaption{lg7\label{fig:skeletonpdhl}}
  \end{minipage}
  \caption{Materialization time in ms for ParallelDHL for the
    ontologies that do not belonging to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.~\label{fig:eval}}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{minipage}{.5\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/1-Finance-rdfox}
    \subcaption{lg2\label{fig:financerdfox}} 
  \end{minipage}
  \begin{minipage}{.5\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/2-molecule-rdfox}
    \subcaption{lg4\label{fig:moleculerdfox}}
  \end{minipage}\\
  \begin{minipage}{.5\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/3-NIF_GrossAnatomy-rdfox}
    \subcaption{lg6\label{fig:grossanatomyrdfox}}
  \end{minipage}
  \begin{minipage}{.5\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/4-skeleton-rdfox}
    \subcaption{lg8\label{fig:skeletonrdfox}}
  \end{minipage}
  \caption{Materialization time in ms for RDFox for the
    ontologies that do not belonging to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.~\label{fig:eval}}
\end{figure}

>>>>

\subsection{The Parallel Tractability of the LUBM Datasets and YAGO Ontologies}

In this part, we analyze two popular datasets, LUBM and YAGO, and show the relations
between these two datasets and the theoretical results of the parallel tractability.

\textbf{LUBM}. In the Semantic Web community, LUBM
(The Lehigh University Benchmark) is proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way.
In the latest version of LUBM,\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
the core ontology contains 48 classes and 32 properties
that are used to describe the departments and the staffs in
universities. By setting different numbers of universities to an ontology-generating
program, users can get datasets of any size based on the core ontology.

In the core ontology of LUBM, the statements about properties, such as inverse property statements,
can be rewritten into the datalog rules of the form (R1), (R2) and (R3) in Table~\ref{tab:dhl}.
Most of the statements about classes can be rewritten into the datalog rules of the form (T1) and (T3)
in Table~\ref{tab:dhl}. Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$ ($\tau$),
where a free variable $y$ is introduced. This kind of axiom is a general case of (T4) in Table~\ref{tab:dhl}
for which $A$ is actually replaced by the top concept $\top$.
Similarly to how we handle (T4), we can also eliminate the free variable $y$
in rule $\tau$ via Skolemization, i.e., replacing the variable $y$ by a new constant $o_R^{A,B}$.
In this way, rule $\tau$ can be rewritten into $A(x)\rightarrow\exists R(x,o_R^{A,B})\wedge B(o_R^{A,B})$.
If we only focus on the materialization task, the rewriting approach via Skolemization guarantees the
completeness and correctness \cite{GrauHKKMMW13}.
On the other hand, rule $\tau$ is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
In summary, if the above kind of rule is not considered,
the materialization of a LUBM dataset can be handled by Algorithm~$\mathsf{A}_{\text{opt}}^{\psi}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet. The latest version
YAGO3 \cite{MahdisoltaniBS15} has more than 10 million entities
(e.g., persons, organizations, cities, etc.)
and contains more than 120 million facts about these entities.

In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called the \emph{YAGO model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}. The YAGO model defines
a set of properties: \texttt{domain, range, subClassOf, subRelationOf} and \texttt{type},
and a set of classes: \texttt{entity, class, relation} and \texttt{acyclicTransitiveRelation}.
The facts in the YAGO model are stated by triples, e.g., ($r_1,\texttt{subRelationOf},r_2$),
which are similar to the RDFS statements.
A group of rules for the reasoning over YAGO ontologies
is specified as follows \cite{SuchanekKW08}:
\begin{enumerate}[leftmargin=8ex,label=(\arabic*),ref=\arabic*]
  \item ($r,\texttt{domain},c$),($x,r,y$)$\rightarrow$($x,\texttt{type},c$)\label{yago:r1}
  \item ($r,\texttt{range},c$),($x,r,y$)$\rightarrow$($y,\texttt{type},c$)\label{yago:r2}
  \item ($c_1,\texttt{subClassOf},c_2$),($x,\texttt{type},c_1$)$\rightarrow$($x,\texttt{type},c_2$)\label{yago:r3}
  \item ($r_1,\texttt{subRelationOf},r_2$),($x,r_1,y$)$\rightarrow$($x,r_2,y$)\label{yago:r4}
  \item ($r,\texttt{type},\texttt{acyclicTransitiveRelation}$),($x,r,y$),($y,r,z$)$\rightarrow$($x,r,z$)\label{yago:r5}
\end{enumerate}

According to the semantics given to YAGO \cite{SuchanekKW08}, the built-in properties in YAGO,
i.e., \texttt{domain, range, subClassOf, subRelationOf} and \texttt{type} act in the same
way as the terms in RDFS statements of the form (1-5) in Table~\ref{tab:rdfs} respectively.
Different from RDFS, YAGO also allows stating acyclic and transitive properties based on the
class \texttt{acyclicTransitiveRelation}; further, any fact in some YAGO ontology cannot be
described by blank nodes. On the other hand,
by carefully observing the rules (\ref{yago:r1}-\ref{yago:r5}) for the reasoning over YAGO ontologies,
one can check that these rules can be rewritten into the datalog rules (T3),\footnote{Both of Rule (\ref{yago:r1}) and
Rule (\ref{yago:r2}) can be rewritten into (T3).} (T1), (R1) and (R3) in Table~\ref{tab:dhl} respectively.
Further, DHL does not restrict any role to be acyclic.
Based on the above analysis, any YAGO ontology can be expressed in DHL
and satisfies the simple-class and the simple-role restrictions. We then have that
for any well-constructed class of YAGO ontologies Algorithm~$\mathsf{A}_{opt}^{\psi}$
can handle all of the ontologies in the class.

In addition to LUBM and YAGO, we further investigate different kinds of ontologies and datasets
including benchmarks, real-world ontologies and datasets that can be expressed in ontology languages.
These ontologies and datsets are collected from the Protege ontology library,\footnote{http://protegewiki.stanford.edu/wiki/Protege$\underline{~}$Ontology$\underline{~}$Library}
Swoogle\footnote{http://swoogle.umbc.edu/} and Oxford ontology lib.\footnote{http://www.cs.ox.ac.uk/isg/ontologies/lib/}
Based on the analysis of these ontologies, we found that, ignoring imports, many of them
belong to $\mathcal{D}_{\textit{\text{dhl}}}$ or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.
All of these investigated ontologies and the analysis results are available online.\footnote{https://github.com/quanzz/PT}

@@

\begin{figure}[htbp]
  \centering
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/1-Finance-simple}
    \subcaption{lg1\label{fig:financepdhl}} 
  \end{minipage}
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/2-molecule-simple}
    \subcaption{lg3\label{fig:moleculepdhl}}
  \end{minipage}\\
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/3-NIF_GrossAnatomy-simple}
    \subcaption{lg5\label{fig:grossanatomypdhl}}
  \end{minipage}
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/4-skeleton-simple}
    \subcaption{lg7\label{fig:skeletonpdhl}}
  \end{minipage}
  \caption{Materialization time in ms for ParallelDHL for the
    ontologies that do not belonging to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.~\label{fig:eval}}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/1-Finance-rdfox}
    \subcaption{lg2\label{fig:financerdfox}} 
  \end{minipage}
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/2-molecule-rdfox}
    \subcaption{lg4\label{fig:moleculerdfox}}
  \end{minipage}\\
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/3-NIF_GrossAnatomy-rdfox}
    \subcaption{lg6\label{fig:grossanatomyrdfox}}
  \end{minipage}
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=\textwidth]{experimentalResults/4-skeleton-rdfox}
    \subcaption{lg8\label{fig:skeletonrdfox}}
  \end{minipage}
  \caption{Materialization time in ms for RDFox for the
    ontologies that do not belonging to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.~\label{fig:eval}}
\end{figure}
